Carbon Nucleosynthesis

### 2023-04-05 thesis/tex/main.tex
thesis updates

\titleformat{\chapter}{\normalfont\normalsize\centering}{Chapter \thechapter:}{1em}{}
\titleformat{\section}{\normalfont\normalsize\centering}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\centering}{\thesubsection}{1em}{}
\titlespacing*{\chapter}{0pt}{0pt}{40pt} \titlespacing*{\section}{0pt}{0pt}{20pt}
\renewcommand{\contentsname}{Table of Contents}
\AtBeginDocument{
      \renewcommand{\bibsection}{\chapter*{\bibname}}
  }

@@

\usepackage{graphicx}	% Including figure files
\graphicspath{{../figures/}}

\usepackage{natbib}
\usepackage[nottoc,numbib]{tocbibind}
% \usepackage[titles]{tocloft} % fig titles
% \renewcommand{\cftfigpresnum}{\figurename\enspace}
% \setlength{\cftfignumwidth}{5em}
% \renewcommand{\cfttabpresnum}{\tablename\enspace}
% \setlength{\cfttabnumwidth}{5em}

@@

As we focus on the yields of carbon, we keep yields for other elements fixed in
most models. Taken from \citet{james+21, james+22}, we let
\begin{itemize}
    \item $Y_\text{O}^\text{CC} = 0.015$
    \item $Y_\text{O}^\text{Ia} = 0$
    \item $Y_\text{Fe}^\text{Ia} = 0.00214?$
    \item $Y_\text{Fe}^\text{CC} = 0.0012$
    \item $Y_\text{N}^\text{CC} = 0.00036$
    \item $y_\text{N}^\text{AGB}(M, Z) = 9\times 10^{-4} M \left(\frac{Z}{Z_\odot}\right)$
\end{itemize}
Also following \citet{james+21, james+22}, we take the SNeIa Delay Time distribution to be a
$t^{-1.1}$ law suggested by the observations of \citet{maoz+12}.

@@

    \caption[AGB DTD]{
    The Delay Time Distribution for AGB STars}

@@

    \caption[SSP AGB carbon]{On the left, we plot the IMF weighted net fractional carbon yields of each AGB model as a function of metalicity.The right panel plots the normalized net carbon AGB yield as a function of time for a population of stars born at the same time at solar metallicity. We note that for the right panel, the minimum of the DTD for V13 is }

@@

In this work, we explore four different sets of AGB star yields representative of the variety of options which also provide reasonably well sampled grids in metallicity and mass necessary for chemical evolution studies.
\begin{itemize}
    \item \citet{cristallo+11} and \citet{cristallo+15} (\cristallo).
    \item \citet{karakas10} (\karakas)
    \item \citet{ventura+13} \ventura
    \item \citet{KL16} \citet{karakas+18} \kl
\end{itemize}
Unless otherwise noted, C11 is the model we choose as our fiducial AGB yield.

@@

Chemical abundances reach an equilibrium. New elements produced in stars balance the elements lost to galactic winds and inside stellar reminants. At the present day, our galaxy has evolved long enough to be well approximated by the chemical equilibriums.

In equilibrium, if we know the yields, than the largest uncertainty is the rate of outflows. This determines the final abundances of evolutionary ... \cite{WAF17}.

@@

\begin{equation}
    y_{\rm C}^{\rm CC} \rightarrow \alpha_{agb}  y_{\rm C}^{\rm CC}
\end{equation}

@@

% y_c^cc = (0.0148 +- 0.00016) Z^{0.154 +- 0.0053}
\begin{equation}
    y_C = (0.0029 \pm 0.0005) + (0.001 \pm 0.0004) \left(\frac{Z}{Z_\odot}\right)
\end{equation}

@@

We only perform this analysis on the C11/15 yields because the range of metallicities of the data (about [M/H] = -0.4, 0.4) only represents about 2 AGB models, relying heavily on our choice of extrapolation. C11/C15 have a finer model grid, allowing us to have 5 theoretical predicted points within this metallicity range, allowing us to be more certain of our conclusion.

@@

To simulate the chemical evolution of a Milky-Way-like galaxy, we use the Versitile Integrator for Chemical Evolution (\VICE\footnote{\VICE~is available at \url{https://github.com/giganano/VICE}}) published in \citet{james+21} and also briefly described in \citet{james+22}.  As a multi-zone model, the galaxy is divided into 200 rings of 0.1kpc, each with separate gas supplies with outflow and inflow rates.

\VICE\ accounts for radial migration by using the results of the \texttt{h277} hydrodynamical simulation. Each VICE SSP is matched to a random nearby star particle \textit{analogue} in \texttt{H277}. VICE approximates the migration of this particle by using a $\sqrt{\text{time}}$-diffusion approximation where the change in galactic radius of a star, $\Delta R$, is $\Delta R \propto \sqrt{\text{t}}$. 
VICE does not account for radial gas flows.

Other approaches have used dynamical arguments to implement their stellar mirgrations. We note that our approach does not introduce more free parameters which may bias conclusions. However, we only consider one dynamic history, and it is still unknown how variations in the dynamic history of a galaxy impact its chemical evolution.

We initially assume an "inside-out" SFH where the star formation rate peaks in the early evolution of the galaxy and is higher closer to the galactic center. The star formation rate, $\Sigma_\star$, for our fiducial model is parameterized by
\begin{equation}
    \dot{\Sigma}_\star \propto \left(1-e^{-t/\tau_{\rm rise}}\right) e^{-t/\tau_{\rm sfh}}
\end{equation}
where $\tau_\text{rise}=2$Gyr is the
and $\tau_{sfh}$ is the rate of decline of the star formation rate, which is based on *****.

@@

VICE normalizes the total mass of star formation against observations from ******.
The gas inflow is calculated based on our star formation parameterization by using a schmit law 
\begin{equation}
    dM_g \propto \dot{\Sigma}^{1/n}
\end{equation}
where we take $n$ to be ****

Salpeter IMF.

@@

\begin{table}
	\centering
    \caption[AGB net solar IMF yields]{For each AGB yield set, our calculation of the effective SFH- and IMF-weighted carbon AGB yield, along with the multiplicative factor to reach an AGB contribution of 20.}
	\label{tab:alpha_agb}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		Study & $y_{\rm C, 0}^{\rm agb}$ & $\alpha_{\rm agb, 20}$\\
		\hline
		C11 & 0.000347 & 2.9\\
		K10 & 0.000585 & 1.7\\
		V13 & 0.000060 & 16.5\\
		K16 & 0.000421 & 2.4\\
		\hline
	\end{tabular}
\end{table}

@@

\section{Validating the APOGEE Subgiant Sample}\label{sec:jack}
Our APOGEE subgiant sample excludes stars marked with the following flags:
\begin{itemize}
\item \texttt{ancillary young embedded cluster}
\item \texttt{ancillary emission line star}
\item \texttt{MIR-deected candidate cluster member}
\item \texttt{selected as part of the EB program}
\item \texttt{selected as part of the young cluster study (in-SYNC)}
\item \texttt{W3/4/5 star forming complex}
\end{itemize}

To select stars only in the subgiant branch, we select stars in a polygon in $\log g$-$T_\text{eff}$ space described by the following cuts:
\begin{equation}
\begin{cases}
\log g \geq 3.5 \\
\log g \leq 0.004\ T_\text{eff} - 15.7 \\
\log g \leq 0.00070588\ T_\text{eff} + 0.358836 \\
\log g \leq -0.0015\ T_\text{eff} + 12.05 \\
\log g \geq 0.0012\ T_\text{eff} - 2.8 \\
\end{cases}
\end{equation}

Instead of excluding stars which have experienced FDU and thus modified carbon abundances as JACK's subgiant sample does, another approach to predict the birth abundances of carbon and nitrogen in stars is to model the predicted abundance effects of dredge up and then correct the measured abundances, as is done in V21.

\newpage
\section{Effects of Alternate AGB Models}\label{sec:alt_agb}

@@

    \caption[Alternate AGB models]{Same as \ref{fig:agb_sims} but comparing our fiducial model to our recommended z-dependent $y_\text{C}^\text{CC}$ model. A plot of the stars of our favored CCSNe model colored by radius. The contours represent the density of V+21 data. The model is given Gaussian scatter equivalent to the mean error in the abundances reported by APOGEE. }
\end{figure}

>>>>

\titleformat{\chapter}{\normalfont\normalsize\centering}{Chapter \thechapter:}{1em}{}
\titleformat{\section}{\normalfont\normalsize\centering}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\centering}{\thesubsection}{1em}{}
% \titlespacing*{\chapter}{0pt}{0pt}{40pt} 
% \titlespacing*{\section}{0pt}{0pt}{20pt}
\renewcommand{\contentsname}{Table of Contents}
\AtBeginDocument{
      \renewcommand{\bibsection}{\chapter*{\bibname}}
  }

@@

\usepackage{graphicx}	% Including figure files
\graphicspath{{../figures/}} \usepackage{natbib}
\usepackage[nottoc,numbib]{tocbibind}
% \usepackage[titles]{tocloft} % fig titles
% \renewcommand{\cftfigpresnum}{\figurename\enspace}
% \setlength{\cftfignumwidth}{5em}
% \renewcommand{\cfttabpresnum}{\tablename\enspace}
% \setlength{\cfttabnumwidth}{5em}

@@

As we focus on the yields of carbon, we keep yields for other elements fixed in
most models. Taken from \citet{james+21, james+22}, we let
\begin{itemize}
    \item $y_\text{Mg}^\text{CC} = 0.00185$
    \item $y_\text{Mg}^\text{Ia} = 0$
    \item $y_\text{Fe}^\text{Ia} = 0.00214$
    \item $y_\text{Fe}^\text{CC} = 0.0012$
    \item $y_\text{N}^\text{CC} = 0.00036$
    \item $\tilde{y}_\text{N}^\text{AGB}(M, Z) = 9\times 10^{-4} M \left(\frac{Z}{Z_\odot}\right)$
\end{itemize}

Also following \citet{james+21, james+22}, we take the SNeIa Delay Time distribution to be a
$t^{-1.1}$ law suggested by the observations of \citet{maoz+12}.

@@

    \caption[AGB yields delayed-time-distribution]{
    The Delay Time Distribution for AGB STars}

@@

    \caption[AGB yield metallicity dependence]{On the left, we plot the IMF weighted net fractional carbon yields of each AGB model as a function of metalicity.The right panel plots the normalized net carbon AGB yield as a function of time for a population of stars born at the same time at solar metallicity. We note that for the right panel, the minimum of the DTD for V13 is }

@@

In this work, we explore four different sets of AGB star yields representative of the variety of options which also provide reasonably well sampled grids in metallicity and mass necessary for chemical evolution studies.
\begin{itemize}
    \item C11: \citet{cristallo+11} and \citet{cristallo+15}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13}
    \item K16: \citet{KL16} \citet{karakas+18}
\end{itemize}
Unless otherwise noted, C11 is the model we choose as our fiducial AGB yield.

@@

Chemical abundances reach an equilibrium. At least with simple enough assumptions. The new elements synthesized in stars balance against elements lost to galactic winds and locked in stellar reminants. While our galaxy is likely not in perfect equilibrium or described by a single, homogenous chemical region, we can still use an equilibrium approximation to understand total yields. 
\cite{WAF17}. \cite{james_dwarf}, \cite{james+22}

Here, we assume a simple one-zone chemical evolution model. A one zone model is essentially zero-dimensional---the gas and stars are considered to be homogenously mixed, so we can represent all quantaties as scalars.
We use $M_{X}$ to represent the mass of element X

@@

\begin{equation}
    \Ycagb \rightarrow \alpha_{agb}  \Ycagb
\end{equation}

@@

% y_c^cc = (0.0148 +- 0.00016) Z^{0.154 +- 0.0053}
\begin{equation}
    \Ycc = y_{\rm C,~0}^{\rm CC} + \zeta \left(\frac{Z}{Z_\odot}\right)
\end{equation}
where we estimate $y_{\rm C,~0}^{\rm CC} = 0.0029\pm 0.0005$ and $\zeta = 0.001\pm0.0004$.

@@

We only perform this analysis on the C11yields because the range of metallicities of the data (about [M/H] = -0.4, 0.4) only represents about 2 AGB models, relying heavily on our choice of extrapolation. C11 have a finer model grid, allowing us to have 5 theoretical predicted points within this metallicity range, allowing us to be more certain of our conclusion.

@@

Unfortunantly, a simple and tidy one-zone model is a poor approximation for our galaxy. The galaxy evolves ``inside out''---where star formation is higher towards the center and in the early universe \citep{bird+13}. Additionally, stars can migrate several kpc over their lifetimes, mixing together different chemical environments across the galaxy \citep{bird+12,sellwood+binney02}. For the rest of this paper, we focus on multizone models, which divide the galaxy into seperate zones and account for stellar migration. 
Our models are created with the Versitile Integrator for Chemical Evolution (\VICE\footnote{\VICE~is available at \url{https://github.com/giganano/VICE}}). \VICE~is published and extensively described in \citet{james+21}. \citet{james+22} include an additional summary. We briefly describe the importaint characteristics of \VICE~here.

As a multi-zone model, the galaxy is divided into 200 rings of 0.1kpc, each with separate gas supplies with outflow and inflow rates. Star formation truncates after 15.5kpc. We initially assume an ``inside-out'' Star Formation History (SFH), where the star formation rate (SFR) $\Sigma_\star$ is given by 
\begin{equation}
    \dot{\Sigma}_\star \propto \left(1-e^{-t/\tau_{\rm rise}}\right) e^{-t/\tau_{\rm sfh}}.
\end{equation}
The SFR reaches a maximum around $\tau_\text{rise}=2$ Gyr, and $\tau{\rm sfh}$ describes the decay timescale of star formation as a function of radius $R$. \VICE~derives $\tau_{\rm sfh}(R)$ from data of MaNGA galaxies in \cite{sanches20}.

@@

The gas inflow is calculated based on an extension of a Kennicutt-Schmidt law (e.g. \cite{kennicutt98}), motivated by 
\begin{equation}
\dot{\Sigma}_{\star} \propto \\
\begin{cases}
    {\Sigma}_{\rm gas} & 2\times 10^7 \leq \Sigma_{\rm gas} \\ 
    {(\Sigma}_{\rm gas})^{3.6} & 5\times 10^6 \leq \Sigma_{\rm gas} < 2\times10^7 \\ 
    {(\Sigma}_{\rm gas})^{1.7} & \Sigma_{\rm gas} < 5\times10^6 \\ 
\end{cases}
\end{equation}

Help? allowing $\tau_\star = \Sigma_{\rm gas}/\dot{\Sigma}_{\star}$, the depletion timescale, to vary as a function of radius and time??

VICE normalizes the total mass of star formation against observations from \cite{LM15}, and uses a double exponential disk profile for the distribution of star formation.  We use a Kroupa IMF \cite{kroupa01}.

\VICE\ accounts for radial migration by using the results of the \texttt{h277} hydrodynamical simulation. Each VICE SSP is matched to a random nearby star particle \textit{analogue} in \texttt{H277}. VICE approximates the migration of this particle by using a $\sqrt{\text{time}}$-diffusion approximation where the change in galactic radius of a star, $\Delta R$, is $\Delta R \propto \sqrt{\text{t}}$. 
VICE does not account for radial gas flows. 
Using the results of a hydrodynamical simulation without modification limits the free parameters in \VICE; however we do not explore different dynamic histories. It is still unknown what role dynamic structure and history plays in chemical evolution.

As the strength of outflows controls the resulting $\alpha$ abundances, \VICE~creates a metallicity gradienty by defining
\begin{equation}
\eta(R) = \frac{y_{\alpha}^{\rm CC}}{Z_{\alpha, \odot}} 10^{(-0.08\text{ kpc}^{-1})(R-4\text{ kpc})+0.3} + r - 1.
\end{equation}
We do explore uniformly scaling yields and outflows in Section REF. While we can constrain relative yields, the absolute scaling of yields and outflows is unknown and chemical evolution models are insensitive to this overal scalling (e.g. \cite{james_dwarf}).

@@

\begin{table}
	\centering
    \caption[AGB net solar IMF yields]{For each AGB yield set, our calculation of the effective SFH- and IMF-weighted carbon AGB yield, along with the multiplicative factor to reach an AGB contribution of 20.}
	\label{tab:alpha_agb}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		AGB Model & $y_{\rm C, 0}^{\rm AGB}$ & $\alpha_{\rm AGB, 20}$\\
		\hline
		C11 & 0.000347 & 2.9\\
		K10 & 0.000585 & 1.7\\
		V13 & 0.000060 & 16.5\\
		K16 & 0.000421 & 2.4\\
		\hline
	\end{tabular}
\end{table}

@@

\section{The APOGEE Subgiant Sample}\label{sec:jack}

We use the criteria outlined in \cite{jack_subgiant} to create a sample of subgiants from APOGEE DR17 \cite{apogee17} as our primary observational constraint. 
\cite{jack_subgiant} selects a region of stars in $\log g$-$T_\text{eff}$ space described by the following polyhedron
\begin{equation}
\begin{cases}
\log g \geq 3.5 \\
\log g \leq 0.004\ T_\text{eff} - 15.7 \\
\log g \leq 0.00070588\ T_\text{eff} + 0.358836 \\
\log g \leq -0.0015\ T_\text{eff} + 12.05 \\
\log g \geq 0.0012\ T_\text{eff} - 2.8 \\
\end{cases}
\end{equation}

This cut isolates a clean sample of subgiants. As First Dredge Up, which affects carbon and nitrogen abundances, only occurs during the ascent onto the red giant branch, subgiant stars are unaffected by this enrichment.

Additionally, we exclued stars in APOGEE marked by the following flags.
\begin{itemize}
\item \texttt{ancillary young embedded cluster}
\item \texttt{ancillary emission line star}
\item \texttt{MIR-deected candidate cluster member}
\item \texttt{selected as part of the EB program}
\item \texttt{selected as part of the young cluster study (in-SYNC)}
\item \texttt{W3/4/5 star forming complex}
\end{itemize}

A second approach in determining birth abundances in stars is to correct the surface abundance effects of FDU, as in \cite{fiorenzo+21}. We do note that there is a slight difference in the yields found through \cite{fiorenzo+21} and \cite{jack_subgiant}. As o

We choose to use \cite{jack_subgiant}'s sample as this does not rely on additional layers of modeling, providing a more direct constraint to our model and limiting our systematic uncertainties.

\newpage
\section{Additional Considerations of Yields}\label{sec:alt_agb}

@@

    \caption[Alternate AGB models]{Comparisons of our 4 different AGB models except with $f_{\rm agb} = 0.2$, and a metallicity dependent CCSNe yield. Both panels plot [C/Mg]-[Mg/Fe], except the left has a bin centered at 0, and the right at [Mg/H]$=0.2$. The higher metallicity bin exposes how many of the AGB models predict downturns near solar metallicity which are not present in our data.}
\end{figure}

\begin{figure}
    \includegraphics{y_mg.pdf}
\end{figure}

\begin{figure}
    \includegraphics{fe_ia.pdf}
\end{figure}
